---
title: "Practical Machine Leaning Course Project"
author: "Camila Vieira"
date: "2023-01-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Importing libraries

This project uses the Random Forest method to predict the classes of the dataset.

```{r}
library(caret)
library(randomForest)
```

### Reading the dataset

```{r}
trainRaw <- read.table("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", sep=",", head=T, row.names=1 ) 
testRaw <- read.table("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", sep=",", head=T, row.names=1 ) 
```

### Cleaning data

Removing all columns with missing values.

```{r}
trainRaw <- trainRaw[, colSums(is.na(trainRaw)) == 0] 
testRaw <- testRaw[, colSums(is.na(testRaw)) == 0] 
```

Maintaining just numeric values in both training and test sets. 

```{r}
classe <- trainRaw$classe
newTrain <- trainRaw[, sapply(trainRaw, is.numeric)]
newTrain$classe <- classe
newTest <- testRaw[, sapply(testRaw, is.numeric)]
```

### Creating data partition

The training data is split into two parts, 70% train and 30% test.

```{r}
set.seed(434)
inTrain <- createDataPartition(newTrain$classe, p=0.70, list=F)
trainData <- newTrain[inTrain, ]
testData <- newTrain[-inTrain, ]
```

### Training and predicting

Crossvalidation with 5 folds is used to detect out of sample error. 

```{r}
controlRF <- trainControl(method="cv", 5)
modelRF <- train(classe ~ ., data=trainData, method="rf", trControl=controlRF, ntree=250)
predictRF <- predict(modelRF, testData)
confusionMatrix(predictRF, as.factor(testData$classe))

result <- predict(modelRF, newTest[, -length(names(newTest))])
result
```


